---
title: "Predicting the default probabilities"
author: "Aylin Mumcular"
date: "29 Mart 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
require(magrittr)
require(tidyverse)
require(knitr)
#require(ISLR)
d <- ISLR::Default
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

## Explore

```{r}
head(d)
dim(d)
summary(d)

d %>% count(default) %>% 
  mutate(prop=n/sum(n))

```

```{r}
d %>% 
  ggplot(aes(balance,income))+
  geom_point(aes(col=default),alpha=0.5)  

d %>% 
  count(default,student) %>% 
  group_by(default) %>%   
  mutate(prop=n/sum(n)) %>% 
  select(-n) %>%   
  spread(student,prop) 

d %>% 
  count(default,student) %>% 
  group_by(student) %>%   
  mutate(prop=n/sum(n)) %>% 
  select(-n) %>%   
  spread(default,prop) #if you know someone is a student what's the likelihoood that this person is gonna default


d %>% ggplot(aes(income))+
  geom_density(aes(fill=student),alpha=0.5) 

d %>% ggplot(aes(balance))+
  geom_density(aes(fill=student),alpha=0.5)  

d %>% ggplot(aes(balance,income))+
  geom_density2d(aes(col=student))

#Students on average make less money than non-students. No relation between income and default. People with high balance tend to default more. 
```


Fit a logistic regression to our default data.

$$
p_i = \frac{e^{x_i^\top \beta}}{1+e^{x_i^\top \beta}}
$$
Let's glm() to find p_i in terms of inputs. 

```{r}
d$default %>% levels()

glmod <- glm(default~student+balance+income,d,family="binomial") 
glmod


```

Test the goodness of the fit of the current model. Using the asymptotic theory:

```{r}
curve(dchisq(x,df=df.residual(glmod)),xlim=c(500,10500)) 
abline(v=deviance(glmod)) 
pvalue <- pchisq(deviance(glmod),
                 df=df.residual(glmod),lower.tail=FALSE)   
pvalue

```

This is overly optimistic, because chi-square approximation to deviance is not good for the logistic regression version of GLM.
Instead let's calculate the null distribution of the deviance. Distribution when the null hypothesis is correct. (Namely, distribution of deviance under $H_0: \text{the current model is correct}$ using bootstrapping.)

```{r}
glmod
repl <- 1000
sim_res <- simulate(glmod,repl) #simulates response. Cooked my own data 
dim(sim_res)

set.seed(545)
deviance_array <- NULL
for(i in seq(repl)){ #get rid of original data, plug cooked ones and refit the model
  d.sim <- mutate(d,default=sim_res[,i]) #replace original ith data
  glmod.sim <- update(glmod,data=d.sim)
  deviance_array[i] <- deviance(glmod.sim)
}

glmod
head(deviance_array)
deviance(glmod)
i=1

```

```{r pval}
deviance_array %>% density %>% plot  #hat distr
abline(v=deviance(glmod)) 

pval <- mean(deviance_array>deviance(glmod)) 
pval 
```
Because the pvalue (`r pval`) is greater than 0.05, we cannot reject the null hypo $H_0: \text{the current model is correct}$ at 5% or less % of level of significance. Our level is fitting fine.

```{r}
predict(glmod,type="response")[1:3]
```

##Important variables?

```{r}
glmod %>% coef

```
Does being a non-student applicant increase the credit default?
Test 

$$

H_0: \beta_\mathrm{StudentYes}=0
$$
I can use **Wald test**. I know that
$\hat\beta_\mathrm{StudentYes}$ has approximately Normal distribution with mean $\beta_\mathrm{StudentYes}$ and $\mathrm{se}(\hat\beta_\mathrm{StudentYes})^2$. This is located at the diagonal entry of

```{r}
vcov(glmod)
```

z values are reported in the summary of the fitted model. 

```{r}
summary(glmod)  #Being a student plays an important role, significant due to p value of 0.006
```

Alternatively, I can test with **likelihood-ratio** test 
```{r}

anova(
  update(glmod,.~.-student-income),
  glmod, test="Chi") 



anova(
  update(glmod,.~.-student), 
  glmod, test="Chi") 

anova(
  update(glmod,.~.-income), 
  #Statistically cannot find sufficient evidence to suggest that income is important. 
  glmod, test="Chi") 
```

**Is the effect of balance equal to 6(10^-3)
H0: Beta_balance=6(10^-3)

Test $H_0:\beta_\mathrm{balance}=0.006$.

```{r}
#Big model logit(pi)=beta0+beta1 studentYes+beta2 balance+beta3 income
#Under H0: logit(pi)=beta0+beta1 studentYes+6(10^-3) balance+beta3 income

small <- glm(default~student+income+offset(6e-3*balance),d,family="binomial")

anova(small,glmod,test="LRT") #LRT is the same with Chi

#we are unable to reject null hypo. That estimate is consistent with the data.
```

Confidence intervals for the effect sizes
```{r}
#confidence interval

confint(glmod, level=0.80)  
```


**Do the effects of balance and income on default risk the same? (beta_balance and beta_income)
H0: Beta_balance=Beta_income=Beta_common

logit pi=beta_0+beta_student studentYes+beta_common (balance+income)

Test $H_0:\beta_\mathrm{balance}=\beta_\mathrm{income}$.

```{r}
small <- glm(default~student+I(balance+income),d,family="binomial") 

d2 <- mutate(d,common=balance+income) #another way

small2 <- glm(default~student+common,d2,family="binomial")

anova(small2,glmod,test="Chi") #Reject and conclude that their effects are not the same.

anova(small,glmod,test="Chi") 
```

```{r}
step(glmod) 

drop1(glmod) 


```

Predictions of default probabilities

```{r}
invisible(predict(glmod)) 

logistic <- function(x) exp(x)/(1+exp(x))
invisible(predict(glmod) %>% logistic)

preds <- predict(glmod,type="response") 


methods("predict")
        
```

The misclassification error is minimized by the Bayes rule which classifies someone as defaulting if predicted default prob şs greater than or equal to 0.5

```{r}
cls <- levels(d$default)[1+(preds>=0.5)] 

```

Confusion table

```{r}
tbl <- table(obs=d$default,preds=cls) 

tbl %>% prop.table %>% diag %>% sum %>% {1-.}
tbl %>% prop.table(1)  #almost 70% of people misclassified. It cannot identify risky people very well. 

table(d$default) %>% prop.table() 


```
 False negative rate (labeling a default as no default) is 228/10000 which is too high. To reduce that I will try to reduce the threshold on the probability of default.
 
```{r}

cls2 <- levels(d$default)[1+(preds>=0.1)] 

tbl2 <- table(obs=d$default,preds=cls2) 

tbl2 %>% prop.table %>% diag %>% sum %>% {1-.}
tbl2 %>% prop.table(1)  
```

Cross-validated (Alternative to AIC, BIC, ANOVA. You can use this method to compare models and decide which variables to keep in the model) misclassification and false negative rates

```{r}
#shuffle the data to avoid data of the same kind. 
nfold <- 10 
set.seed(451)
fold.id <- rep(1:10,len=nrow(d)) %>% sample 
table(fold.id) 
error.rate <- rep(NA,nfold)
fn.rate <- rep(NA,nfold) 



for (i in seq(nfold)) {
  
 
  glmod.train <- glm(default~.,data=d[fold.id!=i,],
                     family="binomial") 
  preds <- predict(glmod.train,newdata=d[fold.id==i,],
                   type="response")
  
  cls <- c("No","Yes")[1+(preds>=.5)] #estimate error rate on unseen data
  
  cnfsn.tbl <- table(obs=d[fold.id==i,"default"],
                     preds=cls)   #confusion table
  error.rate[i] <- prop.table(cnfsn.tbl) %>% diag %>% sum %>% {1-.} #misclassification rate
  
  fn.rate[i] <- prop.table(cnfsn.tbl,1)["Yes","No"] #observed yes predicted no
}
 
m <- error.rate %>% mean
s <- error.rate %>% sd

#95% CI for error rate on the unseen data/generalization error rate

m+c(-1,1)*s*1.96 #alpha=0.05  , m+(c(-1,1))*qnorm(1-alpha/2)*s

m2 <- fn.rate %>% mean
s2 <- fn.rate %>% sd

#95% CI for false negative rate on the unseen data

m2+c(-1,1)*s2*1.96  
```

#Calculate ROC and AUC

```{r,fig.asp=0.5}
glmod
ROC.df <- data_frame(class=d$default,p=fitted(glmod)) %>%  #probabilities
  arrange(desc(p)) %>% #sort this data. Highest p at the top, the one most likely to default
  mutate(NumTP=cumsum(class=="Yes")-(class=="Yes"), 
         NumFP=cumsum(class=="No")-(class=="No"),
         TPR=NumTP/sum(class=="Yes"),
         FPR=NumFP/sum(class=="No"))   #NTP=Number of true positives. Anyone with a threshold of strictly greater than this number will be named as positive. 

plot(TPR~FPR,ROC.df,type="s")
abline(0,1)

#how to calculate area under curve

TPR <- extract2(ROC.df,"TPR")
FPR <- extract2(ROC.df,"FPR")


AUC <- sum((tail(FPR,-1)-head(FPR,-1))*(head(TPR,-1)+tail(TPR,-1))/2) 

AUC


ROC.df %>% mutate(dist=FPR^2+(1-TPR)^2) %>% 
  arrange(dist)





```
























